\documentclass[..\report.tex]{subfiles}

\begin{document}

\section{Future Work}
\label{SEC:FutureWork}

We have now established the importance of a robust and accurately modelled Gulf Stream, however the path to achieving this is not yet clear.
We seek to improve the understanding and thus the representation of the Gulf Stream in moderate-resolution models.
As discussed, there are many processes involved in the dynamics of the ocean and thus there are many different directions to turn to when seeking solutions to this problem. Here we will focus on the interactions and effects of the bathymetry while being careful not to ignore other paths this investigation may lead down. 

\subsection{Identifying Key Processes}
\label{SSEC:IdentifyingKeyProcesses}

%\td{It is speculated *paper*\todo{female author?} that interaction with the bathymetry creates small-scale turbulence and instabilities which can cause bathymetric steering and divert currents. If this is not being represented in coarser models, the energy behind the turbulence must be going elsewhere. \citep{Scott2010} compared simulated kinetic energy over different models against measurements taken and noted that in some areas the total kinetic energy was being held higher in the ocean in the models than observed. 
%It is discrepancies \todo{choose a different word} such as this which could have much wider implications. If the energy is not penetrating to the ocean floor, we cannot expect to be able to replicate the effects of the bathymetry. Perhaps pulling this energy further down (closer to observed values), we would be able to see the energy transfer from \td{the larger eddies} to the turbulence resulting from the bathymetry. }

%\td{As previously discussed there are many processes which could contribute to a Gulf Stream path staying too far South. \citep{Ezer2016b} speculated that amongst other things, the northern branches of the \gls{NRG} would have to be resolved in order to produce an accurate Gulf Stream. \citep{Zhang2007} determined that a significant contribution to the generation of the \gls{NRG} is the bottom vortex stretching resulting from a downslope \gls{DWBC}, which is in turn the result of the interaction with the bathymetry as the \gls{DWBC} crosses the path of the Gulf Stream. Thus we see that this circles back to interaction with the bathymetry.}

The barotropic vorticity diagnostics currently being pursued may lead to some interesting findings into the vorticity balances in different model resolutions and configurations. This will hopefully highlight the processes being resolved by different model configurations. With this insight, we can hope to identify some key areas which could explain the improvements in the representation of the Gulf Stream seen by \citet{Scaife2011a} and others when increasing model resolution.


\subsection{Different Model Formulations}
\label{SSEC:DifferentModelFormulations}

The many different \glspl{OGCM} available allow us to see the effects of many different schemes and parametrisations in use today. In particular the choice of vertical coordinate system has long been discussed, especially in regard to its relevance here.
The representation of the topography and the way it is rendered in the model is a result of the vertical coordinate choice.
\citet{Ezer2016b} revisited the discussion in relation to the Gulf Stream separation and used similar models to determine the most accurate results over different choices for coordinate systems. However, the z-coordinate system chosen did not include any partial or shaved cells.
It is clear that using the z-coordinate system would cause the various slopes in the ocean floor to appear as `steps' which would significantly alter the flow around the area.
It is not so surprising perhaps that in this investigation the models with s-coordinates were found to be the most realistic, given that they allow for a smooth ocean floor.
It is perhaps natural then to question the effect of partial or shaved cells on these findings. As \gls{NEMO} allows for such a diverse range of configurations and schemes, it would be interesting to recreate these results while implementing a z-coordinate system with partial cells to allow for a new take on this review.
The s-coordinate models at coarser resolution provided similar results to z-coordinate models at higher resolution. Implementing our diagnostics on models of varying bathymetrical representation could explain some of these results.
Indeed that has already been mentioned as a possible extension of the barotropic vorticity diagnostics discussed in section \ref{SEC:Diagnostics}.\\

\subsection{Simplified Models}
\label{SSEC:SimplifiedModels}

Simplified models can be a good place to start when investigating theory.  
As discussed previously and shown by \citet{Zhang2007}, the \gls{NRG} and subtropical gyres either side of the Gulf Stream can have significant effects on the path on the Gulf Stream itself, as well as contributing to the strength of it's transport.
To further examine the interplay between the \gls{NRG} and the path of the Gulf Stream, a simple configuration of \gls{NEMO} simulating the double gyre seen in the North Atlantic could be used to examine the dynamics in an idealised model.
Once a more basic state of the set up has been obtained, we could add bathymetrical features mimicking those observed in the North Atlantic and examine the effects of the flow.
This could be investigated using the barotropic vorticity diagnostic or by other means.
Results from these more simplistic configurations could help to influence analysis within the more realistic models. 
As \citet{Tansley2001} used the classic problem of flow past a cylinder to understand the Gulf Stream separation, we too can learn from idealised models and by breaking the problem down to its fundamental aspects.

\subsection{Energy Distribution in the North Atlantic}
\label{SSEC:EnergyDistribution}
  
It has been discussed in section \ref{SEC:DynamicsGulfStream} that a lack of energy transfer, either by lack of interaction with the bathymetry, or via some other process, could be to blame for the low levels of turbulence found in the area surrounding the separation of the Gulf Stream. This is an area which has seen less investigation. Examining the reason for the lack of depth penetration could be an interesting route to go down.
The results from \citet{Tansley2001} showed that more turbulent flow formed jet like streams akin to the Gulf Stream. 
A simplified model could be used here to perhaps artificially pull the energy deeper in the model to allow it to interact with the bathymetry.
The results of this could reveal whether this misplacement of energy is indeed a large contributing factor to the lack of turbulence being represented and if it could be behind the poor representation of the Gulf Stream.
%As seen when \citep{Tansley2001} used a larger Reynolds number, allowing for more turbulent flow, flow passing a quarter-cylinder (akin to the coastline at Cape Hatteras), formed jet like streams with smaller eddies breaking away  from it. This is akin to satellite observations of the Gulf Stream. This leads to the question which many try to answer how can we improve/enhance the turbulence in a model?   

\subsection{Understanding the Evolution of the Gulf Stream}
\label{SSEC:UnderstandingEvolution}
The evolution of the Gulf Stream is of keen interest for climate models.
The change of 30Sv between the pentads 1955-59 and 1970-74 discussed by \citet{Greatbatch1991} is understood to be due to a dramatic change in strength of the subtropical gyre. 
However, global warming brings the threat of a slowdown (or shutdown) of the thermohaline circulation \citep{Gough1998}.
Using \gls{NEMO} to run integrations over longer time periods could provide insight into the evolution of the Gulf Stream and the changing dynamics of the North Atlantic. 
We could use long runs with realistic atmospheric forcing to try to understand the driving forces behind the changes observed. \\

%lans-alpha
\par

The search for better comprehension of the dynamics involved in the global oceans could lead in many directions.
Ulitmately it is only through improved understanding that we can provide a better representation of the Gulf Stream in moderate-resolution climate models. 

\end{document}
